### 2022-04-03 outputs/paper/paper.Rmd
updated Introduction and Data section of paper.rmd. Added in multiple new references in references.bib.

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Tian Yi Zhang
thanks: "Code and data supporting this paper are available at: AOIWDJAWOIDJ"
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
# Load the required packages for this paper
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Introduction

The usage of contraception has become an important part of modern society.
Various contraceptive methods have been invented throughout the years to aid
couples in their family planning efforts. These contraceptive methods range from
hormonal methods and intrauterine methods to barrier methods, providing many
options for both male and female users. Notably, according to the Centers for
Disease Control and Prevention, the barrier method known as condom was the most 
popular male contraceptive used in the United States a few years ago,
with it being around 8.7% of total contraceptives used [@citeCDC].
This is also the case in the country of Kenya, with around 1% of the surveyed
individuals reporting to have used condoms, as denoted in the 1998 DHS
report [@citeDHS].

Another interesting statistic in the studies mentioned above are the high
usage rates of female hormonal contraceptives. 12.6% of contraceptive methods
used in the United States are of the oral hormonal type [@citeCDC], and around
9% of surveyees in Kenya report using oral contraceptives as well [@citeDHS].
This statistic is perplexing in that the usage rate of hormonal contraceptives
are rather high despite the noticeable side effects:
Various studies have suggested in the past that female contraceptives often
lead to high-risk side effects. For example, the U.S. Department of Health &
Human Services notes that oral contraceptives that rely on hormonal controls
can lead to higher blood pressure and more blood clots [@citeDHH]. As such, one
would wonder why male contraceptives do not see more wide usage despite a
relative lack of harmful side effects in comparison to female hormonal
contraceptives.

This paper attempts to analyze potential reasons for the relative lack of
condom usage as a contraceptive method in Kenya utilizing statistics presented
in the 1998 DHS final report on Kenya [@citeDHS]. Specifically, the paper focuses
on finding a, of lack thereof, relationship between the age, geographical
location, and the knowledge of condom(TODO)

WRITE SOMETHING ABOUT FINDINGS; WRITE SOMETHING ABOUT THE STURCTURE OF THE PAPER

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

# Data

The dataset used in this paper is obtained from the PDF "FR102", titled 
"Kenya Demographic and Healthy Survey 1998" and published in 1998 by
the Demographic and Health Surveys Program.^[The PDF can be obtained at
https://dhsprogram.com/publications/publication-fr102-dhs-final-reports.cfm?csSearch=456440_1.]
The raw dataset is parsed using the package `pdftools` [@citePDF] from 
Table 10.14.1 and Table 10.14.2 on page 170 and 171 of the PDF respectively 
[@citeDHS]. The dataset is then cleaned using the `janitor` package 
[@citejani] and the `pointblank` [@citePB] package is used to set up validation
tests for the class and the content of the datasets.
Subsequent graphs and tables shown in this paper also utilize the
`tidyverse` [@citetidy] and `knitr` [@citeknitr] packages.

The PDF where the dataset is parsed from is a report compiled by the DHS program
using the third Demographic and Health Survey conducted in Kenya, the previous
two of which were conducted in 1989 and 1993. [@citeDHS]. According to the
information provided on the DHS Program website, Standard DHS Surveys are
described as surveys with a sample size that consists of 5000 to 30000
households conducted on a five year cycle [@citeDHS2].
Since there is around a five year gap between each consecutive Kenya DHS
survey, we can safely assume that the PDF report containing the raw data is
a standard DHS survey.

The survey used by the DHS Program was distributed in nearly all districts in
Kenya. While the total population of the districts not included in the survey
only amounts to less than four percent of the country's population, it is still
a substantial percentage of the national population that was not sampled from.
The PDF specified that stratified sampling was used, which greatly
contributes greatly to the increase in precision
of the findings from the survey and generally leads to a sample that better
represents the population than a simple random sampling.
However, it was unclear to the readers what
characteristics or attributes were used to divide the population into strata.
Since it was stated later in the PDF that the survey wished to obtain
"district-level data for planning purposed" [@citeDHS], it should be safe to
assume that the strata were constructed based on geographical locations.

The surveys questions were originally constructed fully in English before being
translated into various local languages, and the wording of the questions
were revised several times by local pretest teams before being sent out, 
ensuring that there will not be any inaccuracies or awkward wording present due
to the language barrier.
[@citeDHS]. This is especially important due to the sensitive nature of the
topics asked in the surveys, and helps in making sure that as many
participants answer honestly as possible.

```{r bills, fig.cap="Bills of penguins", echo = FALSE}

```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

```

Talk way more about it.

# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References